Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-08-28 doc/GMP2015/revision_3/s03_drp.tex
Revision 3, close to finished.

@@

\subsection{Canonical DR-Plan}
In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs. Furthermore, we give an efficient \ComplexityCanDRP\ algorithm to find an optimal DR-plan of any independent graph.
% The definition is as follows:

In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ without further specification is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight). Furthermore, we only consider unions and intersections of graphs that are induced subgraphs of a single parent graph $G$. In this case unions and intersections are well defined. For example, the union (resp.\ intersection) of $F_1 = (V_1, E_1)$ and $F_2 = (V_2, E_2)$ is the subgraph of $G$ induced by $V_1\cup V_2$ (resp.\ $V_1\cap V_2$).

% When using intersections and unions of graphs, we are always considering subgraphs of a single parent graph. In this case unions and intersections are well defined. For example, the union of $G_1=(V_1, E_1)$ and $G_2=(V_2, E_2)$ is the subgraph $(V_1\cup V_2, E_1\cup E_2)$; similarly, $G_1\cap G_2$ is the subgraph $(V_1\cap V_2, E_1\cap E_2)$.

@@

    To carry out the induction step, it is sufficient to demonstrate a set of nodes $S$ (of height at most $t$) that must be present in any DR-plan $\drpx{C}$ of graph $C$ that satisfies Property (\ref{def:canonical_drplan:prop1}), including a known optimal one; and furthermore, for any such DR-plan $\drpx{C}$, either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) for all the ancestors $A$ of $S$, $\drpx{A}$ has the minimum possible fan-in of 2 \todo{Shouldn't it be `node $A$ in $\drpx{C}$ has the minimum...'}.

@@

    \medskip\noindent
    \vemph{Claim 1:}
    Let the set of clusters of node $C$ be $C_1,\ldots,C_N$. If the pairwise intersection of clusters is trivial, all of the clusters must be children of $C$ in an optimal DR-plan
    % A set of clusters $C_1,\ldots,C_N$ whose pairwise intersection is trivial must all be children of $C$ in an optimal DR-plan.

@@

    Furthermore, since a canonical DR-plan has nodes with proper rigid \vemph{vertex-maximal} subgraphs as children, if, as in this case, their pairwise intersection is trivial, it follows that any node has at most as many children as a DR-plan without this restriction, because the union of the children must contain all edges of the parent. Therefore, the canonical DR-plan is the optimal choice in this case of trivial intersections.

    \medskip\noindent
    \vemph{Claim 2:}
    Let the set of clusters of node $C$ be $C_1,\ldots,C_N$. If some pair of clusters
    % If some pair in the set of clusters $C_1,\ldots,C_N$ of $C$
    has an isostatic (non-trivial) intersection, then choosing any two as children (minimum possible fan-in) will result in the same maximal antichain of descendants of node $C$.

@@

    % To prove this holds, take the set $[N]=\{1,\dots,N\}$, and denote $I=\bigcap_{k\in [N]}{C_k}$ and $R_k=C\setminus C_k$. Suppose  $C_i$ and $C_j$, where $i\neq j$,  are  the children. For convenience, we will assume all subgraphs are induced subgraphs of $C$. We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$. The isostatic vertex-maximal subgraphs of $C_i$ are $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs. So any two of these are viable children for $C_i$.
    Let $I$ denote the intersection of all the clusters; we call this the \dfn{core}. Let $R_i$ be the graph induced by the edge set of $C$ minus the edge set of $C_i$; we call these the \dfn{appendages}.\footnote{\dfn{Core} and \dfn{appendage} are used in Section~\ref{sec:DRP:algo} and are more formally defined in Definition~\ref{def:seqdrp}.} Note that $C$ is the core plus all appendages, and cluster $C_i$ is the core plus all appendages except $R_i$.
    % To prove this holds,
    Suppose $C_i$ and $C_j$, where $i\neq j$, are taken to be the children of node $C$.
    % We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$.
    The $N-1$ clusters of $C_i$ are the core plus all appendages except $R_i$ and $R_j$, for each $j\neq i$. The pairwise intersection of any of these clusters of $C_i$ will clearly be isostatic, so any two of them are viable children of node $C_i$.
     % $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs. So any two of these are viable children for $C_i$.
    % Since
    % \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
    % the children of $C_i$ will be
    % \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
    % and
    % \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
    % % $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
    % % the children of this node will be
    % % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
    % % and
    % % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
    % for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?}
    Beginning with node $C$, this pattern repeats for $N-1$ levels. Every node in this subtree rooted at $C$ has a fan-in of two (the minimum possible) up through this level. At level $N-1$, we have a set of nodes where each node is the core plus some appendage (with every appendage appearing at least once).
    % This continues for $N-1$ levels, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in [N]$, with every $k$ appearing at least once.
    % At the last level, there are exactly two rigid proper vertex-maximal subgraphs, and hence a unique choice of pair of children.
    Thus, regardless of the sequence of choices of $C_i$ and $C_j$, and of their descendants at each level, the DR-plan has the optimal fan-in of two for every node for $N-1$ levels, and the collection of last level nodes contain the same maximal antichain of subgraphs (for all choices).
    %
    % \medskip\noindent
    % This proof then applies to itself recursively to show that the fan-in of the children will also be minimum.
    %
    % -- say this at top `proof is by induction on the level of the dr=plan'
\end{proof}

This proof of this theorem relies on the following crucial observation and lemma. These will be used again in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper, with modifications to work for other types of qusecs.

@@

        \item\label{lemma:wc_intersection_makes_all_wc}
        If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic. Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.

        \item\label{lemma:uc_intersection_makes_all_uc}
        If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
    \end{enumerate}
\end{lemma*}

@@

The algorithm for finding an optimal DR-plan relies on key structural properties of canonical DR-plans that are revealed by the proof of Theorem~\ref{theorem:canonical_exists_and_is_optimal}. We list these properties before describing the algorithm. Furthermore, in order to deal with the apparent non-uniqueness of a canonical DR-plan (in the case when the children of a node have non-trivial intersections), and to make the canonical DR-plan more malleable for algorithmic purposes, we define a key notion of a \dfn{sequential DR-plan} that is obtained from a canonical DR-plan, and also turns out to be an optimal DR-plan. A sequential DR-plan further clarifies the essential uniqueness of canonical DR-plans. First, we present a recursive definition of the canonical DR-plan, which is more useful in this section, and make an important observation.

\begin{definition}
[Canonical DR-plan]
\label{def:canonical_drp_rec}
    A \dfn{canonical DR-plan} of $G$ is recursively defined as follows:
    \begin{enumerate}
        \item Base case: When $G$ is a single edge, the canonical DR-plan for $G$ is $G$ itself.
        \item In case the pairwise intersections of the proper vertex-maximal rigid subgraphs $C_i$ of $G$  are all trivial, take the children of $G$ to be the roots of the canonical DR-plans for $C_i$.
        \item In case there are two proper vertex-maximal rigid subgraphs $C_i$ and $C_j$ of $G$ with non-trivial intersection, take the children of $G$ to be the roots of the canonical DR-plans for $C_i$ and $C_j$.
    \end{enumerate}
    % When $G$ is a single edge (base case), the canonical DR-plan for $G$ is $G$ itself.
\end{definition}
\todo{What if input is not isostatic? What roots?}

\begin{observation}
[Canonical DR-plan size]
    While all canonical DR-plans of independent graphs are optimal, based on the choice of children in Case (3) above they can have varying \dfn{size} (i.e.\ the number of unique nodes). If we want the canonical DR-plan to have small size (linear \todo{or just polynomial?} in the number of vertices) we can choose the children in the following fashion: In Case (3) above, a canonical DR-plan for $C_i$ (resp.\ $C_j$) is further recursively defined by taking the children of $C_i$ (resp.\ $C_j$) to be the roots of a canonical DR-plan of $C_i\cap C_j$ (which is rigid, by Lemma~\ref{lemma:combined_lemma}), and a canonical DR-plan of $C_i\setminus C_j$ (resp. $C_j\setminus C_i$) (which are not rigid, by Observation~\ref{lemma:union_intersection} and Lemma~\ref{lemma:combined_lemma}).

    Note that a canonical DR-plan can have more than one node representing the same rigid subgraph if and only if Case (3) occurs.
\end{observation}

@@

Now we define the sequential DR-plan in a manner analogous to Definition~\ref{def:canonical_drp_rec}.

\begin{definition}
[(Pseudo-)Sequential DR-plan, appendage, partner, core]
\label{def:seqdrp}
    A \dfn{pseudo-sequential DR-plan} of $G$ is recursively defined as follows:
    \begin{enumerate}
        \item Base case: When $G$ is an edge, the sequential DR-plan for $G$ is $G$ itself.
        \item In case the pairwise intersections of the proper vertex-maximal rigid subgraphs $C_i$ of $G$  are all trivial, take the children of $G$ to be the roots of the sequential DR-plans for $C_i$.
        \item In case there are two proper vertex-maximal rigid subgraphs $C_i$ and $C_j$ of $G$ with non-trivial intersection, take the children of $G$ to be the roots of sequential DR-plans for $C_j\setminus C_i$ (called an \dfn{appendage}), and $C_i$ (called its \dfn{partner}).
        % If the decomposition of the partner $C_i$ results in Case (2), one of the children will be $C_i\cap C_j$, which we call the \dfn{core}.\footnote{Which can be thought of as partner to appendage $C_i\setminus C_j$.}\todo{Is this okay? I don't know how to say this succinctly.}
    \end{enumerate}

    % DR-plans satisfying Requirements 1,2, and 3 are called \dfn{pseudo-sequential.}

    \dfn{Sequential DR-plans} must satisfy another requirement:

>>>>

\newcommand{\pseudosequential}{pseudosequential}
\newcommand{\Pseudosequential}{Pseudosequential}

@@

The goal of this section is to develop an \ComplexityCanDRP\ time complexity algorithm for finding an optimal DR-plan.
To this end, we first introduce a canonical DR-plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs.
While useful for proving optimality, the canonical DR-plan is difficult to work with algorithmically. Therefore, we define the \pseudosequential\ DR-plan which is derived from the canonical and is still optimal.
However, the \pseudosequential\ DR-plan lacks the essential uniqueness of the complete DR-plan. From the \pseudosequential, we derive the sequential DR-plan, which can be found in the same time complexity and is essentially unique.

In this section and in Section~\ref{sec:recomb}, any reference to a graph $G$ without further specification is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight). Furthermore, we only consider unions and intersections of graphs that are induced subgraphs of a single parent graph $G$. In this case unions and intersections are well defined. For example, the union (resp.\ intersection) of $F_1 = (V_1, E_1)$ and $F_2 = (V_2, E_2)$ is the subgraph of $G$ induced by $V_1\cup V_2$ (resp.\ $V_1\cap V_2$).

\subsection{Canonical DR-Plan}

% In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs. Furthermore, we give an efficient \ComplexityCanDRP\ algorithm to find an optimal DR-plan of any independent graph.
% % The definition is as follows:

% In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ without further specification is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight). Furthermore, we only consider unions and intersections of graphs that are induced subgraphs of a single parent graph $G$. In this case unions and intersections are well defined. For example, the union (resp.\ intersection) of $F_1 = (V_1, E_1)$ and $F_2 = (V_2, E_2)$ is the subgraph of $G$ induced by $V_1\cup V_2$ (resp.\ $V_1\cap V_2$).

@@

    To carry out the induction step, it is sufficient to demonstrate a set of nodes $S$ (of height at most $t$) that must be present in any DR-plan $\drpx{C}$ of graph $C$ that satisfies Property (\ref{def:canonical_drplan:prop1}), including a known optimal one; and furthermore, for any such DR-plan $\drpx{C}$, either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) all the ancestors $A$ of $S$ that are descendants of $C$ have the minimum possible fan-in of 2.

@@

    \begin{proof}[Proof of Claim 1]
        Let the set of clusters of node $C$ be $C_1,\ldots,C_N$. If the pairwise intersection of clusters is trivial, all of the clusters must be children of $C$ in an optimal DR-plan
        % A set of clusters $C_1,\ldots,C_N$ whose pairwise intersection is trivial must all be children of $C$ in an optimal DR-plan.

@@

        Furthermore, since a canonical DR-plan has nodes with proper rigid \vemph{vertex-maximal} subgraphs as children, if, as in this case, their pairwise intersection is trivial, it follows that any node has at most as many children as a DR-plan without this restriction, because the union of the children must contain all edges of the parent. Therefore, the canonical DR-plan is the optimal choice in this case of trivial intersections.
    \end{proof}

    % \medskip\noindent
    % \vemph{Claim 1:}

    \begin{proof}[Proof of Claim 2]
        Let the set of clusters of node $C$ be $C_1,\ldots,C_N$. If some pair of clusters
        % If some pair in the set of clusters $C_1,\ldots,C_N$ of $C$
        has an isostatic (non-trivial) intersection, then choosing any two as children (minimum possible fan-in) will result in the same maximal antichain of descendants of node $C$.

@@

        % To prove this holds, take the set $[N]=\{1,\dots,N\}$, and denote $I=\bigcap_{k\in [N]}{C_k}$ and $R_k=C\setminus C_k$. Suppose  $C_i$ and $C_j$, where $i\neq j$,  are  the children. For convenience, we will assume all subgraphs are induced subgraphs of $C$. We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$. The isostatic vertex-maximal subgraphs of $C_i$ are $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs. So any two of these are viable children for $C_i$.
        Let $I$ denote the intersection of all the clusters; we call this the \dfn{core}. Let $R_i$ be the graph induced by the edge set of $C$ minus the edge set of $C_i$; we call these the \dfn{appendages}.\footnote{\dfn{Core} and \dfn{appendage} are used in Section~\ref{sec:DRP:algo} and are more formally defined in Definition~\ref{def:pseudoseqdrp} and~\ref{def:seqdrp}.} Note that $C$ is the core plus all appendages, and cluster $C_i$ is the core plus all appendages except $R_i$.
        % To prove this holds,
        Suppose $C_i$ and $C_j$, where $i\neq j$, are taken to be the children of node $C$.
        % We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$.
        The $N-1$ clusters of $C_i$ are the core plus all appendages except $R_i$ and $R_j$, for each $j\neq i$. The pairwise intersection of any of these clusters of $C_i$ will clearly be isostatic, so any two of them are viable children of node $C_i$.
         % $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs. So any two of these are viable children for $C_i$.
        % Since
        % \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
        % the children of $C_i$ will be
        % \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
        % and
        % \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
        % % $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
        % % the children of this node will be
        % % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
        % % and
        % % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
        % for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?}
        Beginning with node $C$, this pattern repeats for $N-1$ levels. Every node in this subtree rooted at $C$ has a fan-in of two (the minimum possible) up through this level. At level $N-1$, we have a set of nodes where each node is the core plus some appendage (with every appendage appearing at least once).
        % This continues for $N-1$ levels, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in [N]$, with every $k$ appearing at least once.
        % At the last level, there are exactly two rigid proper vertex-maximal subgraphs, and hence a unique choice of pair of children.
        Thus, regardless of the sequence of choices of $C_i$ and $C_j$, and of their descendants at each level, the DR-plan has the optimal fan-in of two for every node for $N-1$ levels, and the collection of last level nodes contain the same maximal antichain of subgraphs (for all choices).
    \end{proof}
    %
    % \medskip\noindent
    % \vemph{Claim 2:}
    %
    % \medskip\noindent
    % This proof then applies to itself recursively to show that the fan-in of the children will also be minimum.
    %
    % -- say this at top `proof is by induction on the level of the dr=plan'
    % The proofs of Claim 1 and Claim 2 prove the theorem.
\end{proof}

This proof of the theorem relies on the following crucial observation and lemma. These will be used again in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper, with modifications to work for other types of qusecs.

@@

        \item\label{lemma:wc_intersection_makes_all_wc}
        If $C_i\cup C_j$ is isostatic, then for all $k$, $C_i\cup C_k$ is isostatic. Alternatively, if $C_i\cup C_j=C$, then for all $k$, $C_i\cup C_k=C$.

        \item\label{lemma:uc_intersection_makes_all_uc}
        If $C_i\cap C_j$ is trivial, then for all $k$, $C_i\cap C_k$ is trivial.
    \end{enumerate}
\end{lemma*}

@@

% The algorithm for finding an optimal DR-plan relies on key structural properties of canonical DR-plans that are revealed by the proof of Theorem~\ref{theorem:canonical_exists_and_is_optimal}. We list these properties before describing the algorithm. Furthermore, in order to deal with the apparent non-uniqueness of a canonical DR-plan (in the case when the children of a node have non-trivial intersections), and to make the canonical DR-plan more malleable for algorithmic purposes, we define a key notion of a \dfn{sequential DR-plan} that is obtained from a canonical DR-plan, and also turns out to be an optimal DR-plan. A sequential DR-plan further clarifies the essential uniqueness of canonical DR-plans. First, we present a recursive definition of the canonical DR-plan, which is more useful in this section, and make an important observation.

The algorithm for finding an optimal DR-plan relies on key structural properties of canonical DR-plans that are revealed by the proof of Theorem~\ref{theorem:canonical_exists_and_is_optimal}. We begin by redefining the canonical DR-plan in a recursive manner and also by recursively defining the new \dfn{\pseudosequential\ DR-plan} to highlight the similarities. We show that this \pseudosequential\ DR-plan is optimal (i.e.\ has the smallest maximum fan-in over all nodes) and has at most the same overall size (i.e.\ number of unique nodes) as a canonical DR-plan, and is in general smaller.
% By taming the large number of nodes with redundant subgraphs (in the case when the children of a node have non-trivial intersections),
By ensuring that no two nodes contain the same subgraph,
the \pseudosequential\ DR-plan is more malleable for algorithmic purposes.

\begin{definition}
[Canonical DR-plan]
\label{def:canonical_drp_rec}
    A \dfn{canonical DR-plan} of an isostatic $G$ is recursively defined as follows:
    \begin{enumerate}
        \item Base case: When $G$ is a single edge, the canonical DR-plan for $G$ is $G$ itself.
        \item In case the pairwise intersections of the proper vertex-maximal rigid subgraphs $C_i$ of $G$  are all trivial, take the children of $G$ to be the roots of the canonical DR-plans for $C_i$.
        \item In case there are two proper vertex-maximal rigid subgraphs $C_i$ and $C_j$ of $G$ with non-trivial intersection, take the children of $G$ to be the roots of the canonical DR-plans for $C_i$ and $C_j$.
    \end{enumerate}
    % When $G$ is a single edge (base case), the canonical DR-plan for $G$ is $G$ itself.
\end{definition}
% \todo{What if input is not isostatic? What roots?}

% \begin{observation}
% [Canonical DR-plan size]
%     While all canonical DR-plans of independent graphs are optimal, based on the choice of children in Case (3) above they can have varying \dfn{size} (i.e.\ the number of unique nodes). If we want the canonical DR-plan to have small size (linear \todo{or just polynomial?} in the number of vertices) we can choose the children in the following fashion: In Case (3) above, a canonical DR-plan for $C_i$ (resp.\ $C_j$) is further recursively defined by taking the children of $C_i$ (resp.\ $C_j$) to be the roots of a canonical DR-plan of $C_i\cap C_j$ (which is rigid, by Lemma~\ref{lemma:combined_lemma}), and a canonical DR-plan of $C_i\setminus C_j$ (resp. $C_j\setminus C_i$) (which are not rigid, by Observation~\ref{lemma:union_intersection} and Lemma~\ref{lemma:combined_lemma}).

%     Note that a canonical DR-plan can have more than one node representing the same rigid subgraph if and only if Case (3) occurs.
% \end{observation}

@@

Now we define the \pseudosequential\ DR-plan in a manner analogous to Definition~\ref{def:canonical_drp_rec}.

\begin{definition}
[\Pseudosequential\ DR-plan, appendage, partner]
\label{def:pseudoseqdrp}
    A \dfn{\pseudosequential\ DR-plan} of an isostatic $G$ is recursively defined as follows:
    \begin{enumerate}
        \item Base case: When $G$ is an edge, the \pseudosequential\ DR-plan for $G$ is $G$ itself.
        \item In case the pairwise intersections of the proper vertex-maximal rigid subgraphs $C_i$ of $G$  are all trivial, take the children of $G$ to be the roots of the \pseudosequential\ DR-plans for $C_i$.
        \item In case there are two proper vertex-maximal rigid subgraphs $C_i$ and $C_j$ of $G$ with non-trivial intersection, take the children of $G$ to be the roots of \pseudosequential\ DR-plans for $C_j\setminus C_i$ (called an \dfn{appendage}), and $C_i$ (called its \dfn{partner}).
    \end{enumerate}
\end{definition}

\begin{remark}
[Size of \pseudosequential\ DR-plan]
    Any rigid subgraph appears in at most one node of a \pseudosequential\ DR-plan. Also, if an edge $e$ of $G$ belongs in the rigid subgraph at any two nodes $A$ and $B$ of a \pseudosequential\ DR-plan, then either $A\subset B$ or $B\subset A$. Since the leaves of \pseudosequential\ DR-plan tree are the $O(|V(G)|)$ edges of the independent input graph, this implies the size of the DR-plan is $O(|V(G)|)$.
\end{remark}

\begin{observation}
[Sequential is optimal]
    Any \pseudosequential\ DR-plan of an independent graph is optimal.
\end{observation}

\begin{proof}
    We need to show the statement:
    (*) the max fan-in (i.e.\ number of children) of any \pseudosequential\ DR-plan for an independent graph $G$ is no larger that of some canonical DR-plan for $G$. By Theorem~\ref{theorem:canonical_is_optimal}, this would imply that the \pseudosequential\ DR-plan is optimal.

    Recall that both \pseudosequential\ and canonical DR-plans are defined recursively. We will prove the statement (*) by induction on the height of a \pseudosequential\ DR-plan for $G$. The base case (height of 0, i.e.\ single edges) trivially holds. Induction hypothesis: (*) holds for independent graphs $G$ with \pseudosequential\ DR-plans of height $h$.

    Induction step: a \pseudosequential\ DR-plan of height $h+1$ rooted at a node $G$ consists of \pseudosequential\ DR-plans (of height at most $h$) rooted at the children of $G$, which we call the set $C$. It is sufficient to show (a) the children $C$ will exist somewhere in some canonical DR-plan, and (b) $|C|$ is less than or equal to the max fan-in of some canonical DR-plan. Given (a), then by the recursive definition of canonical DR-plans, these nodes are the roots of canonical DR-plans, thus the induction hypothesis applies to the nodes in $C$. Additionally given (b), the proof of the induction step of (*) is complete.

    Case (1):
    When the isostatic vertex-maximal proper subgraphs of $G$ have trivial intersections, the children of $G$ in any canonical DR-plan are the same set $C$ of children in any \pseudosequential\ DR-plan. Thus both conditions (a) and (b) are immediately satisfied.

    Case (2):
    When the isostatic vertex-maximal proper subgraphs of $G$ have non-trivial intersections, the children $C$ are the isostatic components of some appendage, $A_i$, and its partner. The partner will be a child of $G$ in some canonical DR-plan. Furthermore, in any canonical DR-plan there will be some node $N-1$ levels below $G$ (where $N$ is the number of appendages) containing the core (the common intersection of all the isostatic vertex-maximal proper subgraphs of $G$) plus appendage $A_i$. In some canonical DR-plan, the children of this node will be the components of $A_i$ and the core. This node has the same fan-in as $G$ in the \pseudosequential\ DR-plan, satisfying (b), and, along with the partner (a child of $G$), shows that some canonical DR-plan has all the nodes in $C$, thereby satisfying (a).
\end{proof}

\begin{definition}
[\Branch]
    Given a \pseudosequential\ DR-plan $P_G$ for an isostatic graph $G$, and an edge $e \in G$, the \branchGePG\ is the subtree of the DR-plan consisting of the path from the leaf containing $e$ to the root containing $G$, together with the children of all the nodes on this path, which are the leaves of \branchGePG. See Figure~\ref{fig:seq_proof} for examples.
\end{definition}

\begin{observation}
[\Pseudosequential\ DR-plan recursively from \branches]
\label{obs:seqplan_rec_from_branches}
    A \pseudosequential\ DR-plan $P_G$ for an isostatic graph $G$ is obtained from \branchGePG\ by recursively attaching to each of its leaves $L$ a \pseudosequential\ DR-plan $P_L$ for $L$.
\end{observation}

The next two Lemmas are the crux of our $O(|V|^3)$ algorithm to find a \pseudosequential\ DR-plan.

\begin{lemma}
[\Branch\ leaves from components]
    Let $G$ be isostatic, $e$ be an edge in $G$, and $\componentsx{G\setminus e}$ be the set of maximal rigid components of $G\setminus e$. Then there is a \pseudosequential\ DR-plan $P_G$ for $G$ such that $\componentsx{G\setminus e}$ is exactly the set of leaves of \branchGePG\ (minus $e$ itself).
\end{lemma}

\begin{proof}
    Follows from the structure of a \pseudosequential\ DR-plan and the definition of \branchGePG.
\end{proof}

% \ClearMyMinHeight
% \SetMyMinHeight{0.195}{../../img/svg/seq_proof_figs_a}
% \SetMyMinHeight{0.195}{../../img/svg/seq_proof_figs_b}
% \SetMyMinHeight{0.195}{../../img/svg/seq_proof_figs_c}
% \SetMyMinHeight{0.195}{../../img/svg/seq_proof_figs_d}
% \SetMyMinHeight{0.195}{../../img/svg/seq_proof_figs_e}
\ClearMyMinHeight
\SetMyMinHeight{0.3}{../../img/svg/seq_proof_figs_a}
\SetMyMinHeight{0.3}{../../img/svg/seq_proof_figs_b}
\SetMyMinHeight{0.3}{../../img/svg/seq_proof_figs_c}
\SetMyMinHeight{0.3}{../../img/svg/seq_proof_figs_d}
\SetMyMinHeight{0.3}{../../img/svg/seq_proof_figs_e}

\begin{figure*}\centering%
    % \begin{subfigure}[t]{0.27\linewidth}
    %     \includegraphics[width=\linewidth]{../../img/svg/seq_proof_figs_a}
    %     \caption{}\label{fig:seq_proof:a}
    % \end{subfigure}%

    % \begin{subfigure}{0.195\linewidth}\centering
    %     \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_a}
    %     \caption{}\label{fig:seq_proof:a}
    % \end{subfigure}%
    % %
    % \hfill
    % \begin{subfigure}{0.195\linewidth}\centering
    %     \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_b}
    %     \caption{}\label{fig:seq_proof:b}
    % \end{subfigure}%
    % %
    % \begin{subfigure}{0.195\linewidth}\centering
    %     \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_c}
    %     \caption{}\label{fig:seq_proof:c}
    % \end{subfigure}%
    % %
    % \hfill
    % \begin{subfigure}{0.195\linewidth}\centering
    %     \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_d}
    %     \caption{}\label{fig:seq_proof:d}
    % \end{subfigure}%
    % %
    % \hfill
    % \begin{subfigure}{0.195\linewidth}\centering
    %     \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_e}
    %     \caption{}\label{fig:seq_proof:e}
    % \end{subfigure}%

    \begin{subfigure}{0.3\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_a}
        \caption{}\label{fig:seq_proof:a}
    \end{subfigure}%
    %
    \hfill
    \begin{subfigure}{0.3\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_b}
        \caption{}\label{fig:seq_proof:b}
    \end{subfigure}%
    %
    \hfill
    \begin{subfigure}{0.3\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_c}
        \caption{}\label{fig:seq_proof:c}
    \end{subfigure}%

    % \hspace{1cm}
    \begin{subfigure}{0.3\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_d}
        \caption{}\label{fig:seq_proof:d}
    \end{subfigure}%
    %
    \hspace{2cm}
    \begin{subfigure}{0.3\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/seq_proof_figs_e}
        \caption{}\label{fig:seq_proof:e}
    \end{subfigure}%
    % \hspace{1cm}

    \caption{
    This figure illustrates the notion of a \branch, how it relates to the components of $G\setminus e$, and the variety of cases discussed in the proof of Lemma~\ref{lem:branch-from-branch-leaves}. Each subfigure is $\branchx{G,e,P_G}$ for some edge $e$ (resp.\ $\branchx{G,f,P_G}$), the white nodes are the path from $G$ to $e$ (resp.\ $f$), and the black nodes are the set $\componentsx{G\setminus e}$ (resp.\ $G\setminus f$). In subfigures (\ref{fig:seq_proof:b}), (\ref{fig:seq_proof:c}), and (\ref{fig:seq_proof:d}), nodes are labeled with the case they fall under in the proof of Lemma~\ref{lem:branch-from-branch-leaves} if they are chosen to be $D$ (given the choice of $L$ discussed below), any unlabeled node falls under Case (1b).
    %
    Figure (\ref{fig:seq_proof:a}) has nodes labeled $L_1$ and $L_2$ ($L_1'$ will be discussed later), from which different edges $f$ are taken in subsequent figures. Note that the node labeled `Core+$A_1$+$A_2$' is a subgraph of $G$ that has 2 isostatic vertex-maximal proper subgraphs intersecting non-trivially. The core is the intersection of these subgraphs and $A_i$ is an appendage.
    %
    Figure (\ref{fig:seq_proof:b}) uses $f=L_1$.
    %
    Figure (\ref{fig:seq_proof:c}) uses an arbitrary $f\in L_2$.
    %
    Figure (\ref{fig:seq_proof:d}) uses an $f$ in the core in $L_3$.
    % Note that based on the choice of $e$ as the first edge, we get a \pseudosequential\ DR-plan that looks like Figure (\ref{fig:seq_proof:c}), if we had chosen an edge from appendage $A_2$ the second and third level would be flipped. \todo{Shows cases...}
    %
    Figure (\ref{fig:seq_proof:e}) uses an $f$ in the appendage $A_1$ in $L_3$. Note that this contains a node $D$ that has Case (3). With this $D$, $D\cup L_3$ is `Core+$A_1$+$A_2$' and $D\cap L_3$ is `Core'. The leaves labeled $L_3'$ in Figure (\ref{fig:seq_proof:a}) are the leaves that will allow us to find the node on the path from $G$ to $e$ that is their sibling (the parent of $e$).
    %
    % Figure (\ref{fig:seq_proof:a}) depicts $\branchx{G,e,P_G}$, the black nodes are the $\componentsx{G\setminus e}$. The node labeled `Core+$A_1$+$A_2$' is a subgraph of $G$ that has 2 isostatic vertex-maximal proper subgraphs intersecting non-trivially.
    %
    % Figure (\ref{fig:seq_proof:b}) depicts $\branchx{G,f,P_G}$ of the edge labeled $f\in L_2$ (from Figure (\ref{fig:seq_proof:a})). Similarly, the black nodes are the $\componentsx{G\setminus f}$. These leaves are labeled with the cases discussed in the proof of Lemma~\ref{lem:branch-from-branch-leaves}. \todo{emphasize in beginning.}
    %
